\section*{Research Question}

How do mandatory paternity leave quotas affect the life-cycle gender wage gap and household welfare? Specifically, does forcing fathers to take a minimum share of parental leave reduce the long-run earnings gap between partners, and at what welfare cost?

\section*{Motivation}

Using Danish register data, \citet{kleven2019children} document a sharp ``child penalty'': mothers' earnings fall by roughly 20 percentage points relative to childless women after first birth, while fathers are unaffected. A central driver is that mothers take approximately 80 percent of total parental leave, interrupting their human capital accumulation. \citet{kleven2019childpenalties} show that this pattern is pervasive across countries, and \citet{patnaik2019reserving} finds that earmarked paternity quotas substantially shift leave allocation toward fathers. Whether such quotas reduce the gender wage gap, and whether the long-run gain in female earnings outweighs the short-run household income loss, is a quantitative question that requires a structural life-cycle model.

\section*{Model}

We propose a finite-horizon two-earner household model in the spirit of \citet{borella2023marriage}, indexed by $i \in \{1,2\}$ (father and mother), over $T$ periods. The household maximises joint expected lifetime utility by jointly choosing consumption, hours, and the leave allocation, subject to wages that depend on accumulated human capital and a policy constraint on the father's minimum leave share. Table~\ref{tab:params} lists the parameters of the model.

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}llc@{}}
\toprule
\textbf{Symbol} & \textbf{Description} & \textbf{Source} \\
\midrule
$T$            & Number of life-cycle periods                                & Calibrated \\
$\rho$         & Discount factor                                              & Calibrated \\
$r$            & Risk-free interest rate                                      & Calibrated \\
$\eta$         & CRRA coefficient (consumption curvature)                     & Estimated (SMM) \\
$\gamma$       & Curvature on labor disutility                                & Estimated (SMM) \\
$\beta_0$      & Labor disutility weight, no child                            & Estimated (SMM) \\
$\beta_1$      & Additional labor disutility when a child is present          & Estimated (SMM) \\
$\bar{w}$      & Pre-tax wage level                                           & Calibrated \\
$\tau$         & Labor income tax rate                                        & Legislation \\
$\alpha$       & Return on human capital                                      & Estimated (SMM) \\
$\delta_1$     & Human capital depreciation during leave, father              & Estimated (SMM) \\
$\delta_2$     & Human capital depreciation during leave, mother              & Estimated (SMM) \\
$p_b$          & Per-period birth probability when $n_t = 0$                  & Calibrated \\
$\bar{L}$      & Total parental leave time per birth                          & Calibrated \\
$\phi$         & Wage replacement rate during leave (\emph{barselsdagpenge})  & Legislation \\
$\bar{q}$      & Mandatory paternity leave quota (policy instrument)          & Counterfactual \\
\bottomrule
\end{tabular}
\caption{Model parameters and their role.}
\label{tab:params}
\end{table}

\subsection*{Recursive Formulation}

Let the state be $s_t = (a_t, K_{1,t}, K_{2,t}, n_t)$, where $a_t$ is assets, $K_{i,t}$ is human capital, and $n_t \in \{0,1\}$ indicates the presence of a child. The Bellman equation is
\begin{equation}
    V_t(s_t) = \max_{c_t,\, h_{1,t},\, h_{2,t},\, \lambda_t}\;\Bigl\{ U(c_t, h_{1,t}, h_{2,t}, n_t) + \rho\,\mathbb{E}_t\!\left[V_{t+1}(s_{t+1})\right]\Bigr\},
\end{equation}
with period utility
\begin{equation}
    U(c_t, h_{1,t}, h_{2,t}, n_t) = \frac{c_t^{1+\eta}}{1+\eta} - (\beta_0 + \beta_1 n_t)\!\left[\frac{h_{1,t}^{1+\gamma}}{1+\gamma} + \frac{h_{2,t}^{1+\gamma}}{1+\gamma}\right],
\end{equation}
where $\beta_1 > 0$ captures the additional disutility of market work in the presence of a child.

\subsection*{Wages and Human Capital}

The after-tax wage of agent $i$ is
\begin{equation}
    w_{i,t} = (1-\tau)\,\bar{w}\,(1 + \alpha\, K_{i,t}),
\end{equation}
and human capital evolves as
\begin{equation}
    K_{i,t+1} = K_{i,t}\,(1 - \delta_i\, l_{i,t}) + h_{i,t},
\end{equation}
where $l_{i,t} \in [0,1]$ is the share of the period spent on leave and $\delta_i \geq 0$ is the depreciation rate during leave. The key asymmetry is $\delta_2 \geq \delta_1$: leave is more costly for the mother's career, consistent with the child-penalty evidence in \citet{kleven2019children}.

\subsection*{Fertility, Leave, and Policy}

A child arrives with probability $p_b$ if $n_t = 0$. Upon birth, total leave $\bar{L}$ is split between partners: the father takes share $\lambda_t$ and the mother takes $1-\lambda_t$, with each receiving a wage replacement at rate $\phi$. Each agent's time is split between work and leave,
\begin{equation}
    h_{i,t} + l_{i,t} \;\leq\; 1, \qquad l_{i,t} \geq 0,\; h_{i,t} \geq 0,
\end{equation}
and the mandatory paternity quota imposes
\begin{equation}
    \lambda_t \geq \bar{q}\quad\text{if a birth occurs in period }t,
\end{equation}
where $\bar{q} \in [0, 0.5]$ is the policy instrument. The household budget constraint is
\begin{equation}
    c_t + a_{t+1} = (1+r)\,a_t + \sum_{i=1}^{2}\bigl[w_{i,t}\,h_{i,t} + \phi\,\bar{w}\,l_{i,t}\bigr].
\end{equation}

\subsection*{Mechanism}

Two forces interact. First, a higher $\bar{q}$ shifts leave from mother to father, dampening the depreciation $\delta_2 (1-\lambda_t)\bar{L}$ on her human capital and raising her future wage through (3)--(4). Second, because $\delta_1 < \delta_2$, reallocating leave is costly in terms of \emph{aggregate} human capital and household income. The optimal $\bar{q}$ balances these gains and losses against the policy-maker's distributional objective. The model is solved by backward induction.

\section*{Identification and Counterfactuals}

We use the model to answer the research question in two steps. First, we estimate the structural parameters $\theta = (\eta, \gamma, \beta_0, \beta_1, \alpha, \delta_1, \delta_2, p_b, \rho)$ by simulated method of moments, matching pre-birth and post-birth profiles of hours, wages, and leave take-up by gender. The asymmetry $\delta_2 - \delta_1$ is identified by the cross-gender difference in the post-birth wage trajectory documented by \citet{kleven2019children}.

Second, we run counterfactual simulations across $\bar{q} \in \{0,\, 0.10,\, 0.25,\, 0.50\}$ and report (i) the life-cycle gender wage gap $G_t = (w_{1,t}-w_{2,t})/w_{1,t}$, (ii) female labor supply, and (iii) the consumption-equivalent welfare gain $\xi(\bar{q})$, defined as the proportional increase in lifetime consumption under $\bar{q}=0$ that would leave the household indifferent to the quota policy:
\begin{equation}
    V_0\!\left(s_0;\, \bar{q}\right) \;=\; V_0\!\left(s_0;\, \bar{q}=0,\, c_t \cdot (1+\xi)\right).
\end{equation}
Tracing $\xi(\bar{q})$ and $G_t(\bar{q})$ jointly maps out the policy frontier between long-run wage convergence and short-run household income loss.

\section*{Data}

To estimate the structural parameters and validate the counterfactuals, we use:
\begin{itemize}
    \item \textbf{Danish administrative registers (IDA):} matched employer-employee data on wages and hours by gender and parental status, used to identify $\alpha$, $\delta_1$, $\delta_2$ and to match the child penalty in \citet{kleven2019children}.
    \item \textbf{Barselstatistik (Statistics Denmark):} parental leave spell lengths by parent, used to calibrate $\bar{L}$ and the baseline distribution of $\lambda_t$.
    \item \textbf{Danish Labor Force Survey (AKU):} hours worked and employment rates by parental status, used to pin down the labor disutility parameters $\beta_0$, $\beta_1$, $\gamma$.
    \item \textbf{Danish social-insurance rules:} the statutory replacement rate $\phi \approx 0.9$ (\emph{barselsdagpenge}) is taken directly from legislation and requires no estimation.
\end{itemize}

\noindent For external validation, the introduction of the 2022 Danish earmarked paternity quota provides a quasi-experimental setting against which the model's predicted change in $\lambda_t$ and $G_t$ can be benchmarked, similar in spirit to \citet{patnaik2019reserving} and \citet{carneiro2015flying}.
